Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating

### 2017-09-15 SubGenome Analysis.Rmd
Load, filter, and plot isoform frequencies

@@

```{r diffExpression_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
## Attach log2(FPKM_mean) values to homeologous pairs
expressedPairs <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  inner_join(expressedGenes, by=c("Maize1"="geneID")) %>%
  rename(FPKM_mean1=FPKM_mean) %>%
  inner_join(expressedGenes, by=c("Maize2"="geneID")) %>%
  rename(FPKM_mean2=FPKM_mean)

# Expression Diffs
grid.newpage()
g <- draw.pairwise.venn(
  area1 = expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2) %>% nrow(),
  area2 = expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2) %>% nrow(),
  cross.area = intersect(expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2),
                         expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2)) %>%
                          nrow(), 
  category = c("Sub1", "Sub2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="Genes with 2-fold higher\nexpression by subgenome")
```

>>>>

In order to see if one subgenome tends to have more isoforms, we plot the density of isoform counts from each subgenome based on isoforms in the v4 release 32 GFF file. We scale the x axis using log10, as we have a very long tail on this data (i.e. most counts are 1, but a few range to nearly 600). The densities track each other very strongly, so there isn't likely anything to look for here.

```{r subgenome_isoform_counts, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=6, fig.height=4, fig.align='center'}
## Plot frequency of isoform counts for each gene
data <- geneTranscript.counts %>% inner_join(subgenome, by=c("gene"="gene2")) %>% select(gene, n, subgenome) %>% filter(subgenome=="sub1" | subgenome=="sub2")
ggplot(data, aes(n,color=subgenome)) + 
  geom_density() + 
  scale_x_log10() +
  labs(y = "Density (Isoform Counts)", 
       x = "log10(Isoform Counts)", 
       title = "Comparison of Density of Isoform Counts between\nGenes in Subgenome 1 and Subgenome 2"
  )
```

@@

```{r diffExpression_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
# Expression Diffs
grid.newpage()
g <- draw.pairwise.venn(
  area1 = expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2) %>% nrow(),
  area2 = expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2) %>% nrow(),
  cross.area = intersect(expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2),
                         expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2)) %>%
                          nrow(), 
  category = c("Sub1", "Sub2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="Genes with 2-fold higher\nexpression by subgenome")
```
